---
title: "Stat Comp HW 4"
author: "Michael Ackerman"
date: "October 27, 2015"
output: html_document
---

Secant Algorithm Code:
```{r}
seca <- function(x, f, tol=1e-10, maxit=1000){
  i<-2
  while( abs(f(x[i])) > tol & i < maxit){
    x[i+1] <- x[i]-f(x[i])*((x[i]-x[i-1])/(f(x[i])-f(x[i-1])))
    i <- i+1
  }
    if(i==maxit) stop('pick a bigger number, bro')
    x[i]
}

## Initial guesses in vector form 
x<- c(-8,-7)
## Validation Function
f <- function(x) { cos(x)-x }
## Call Code
seca(x,f)

## Newton Method
newton <- function(guess, f, df, tol=1e-10, maxit=1000){
  i<-1
  while( abs(f(guess)) > tol & i < maxit){
    guess <- guess - f(guess)/df(guess)
    i <- i+1
  }
    if(i==maxit) stop('method didnt converge')
    guess
}
## Validation Function derivative
df <- function(x) { -sin(x)-1}

## Time comparison
system.time(replicate(10000, newton(-8,f,df))) - system.time(replicate(10000, seca(x,f)))
```

For this time comparison, I ran the above code many times and recieved an even amount of positive and negative values for "user time". 

**2)**
This code will play Craps 'n' number of times.
```{r}
Craps <- function(games=1, seed= .Random.seed){
  set.seed(seed) 
  results <- c()
  for(i in 1:games){ 
die <- sum(ceiling(6*runif(2)))
win <- -1
while(win==-1){
   if(die == 7 | die == 11) {win <- 1}
   die2 <- sum(ceiling(6*runif(2)))
   if(die2 == 7 | die2 == 11) {win <- 0}
   else if(die==die2){win <- 1}
 } 
results <- c(results, win)
}
return(results)
} 

Craps(games= 3,seed = 100)
```

In the code, "0" denotes a loss. So with a seed(100), we'll lose the first three games.

This code finds the seed value that will produce 10 straight wins.

```{r}
i <- 1
set.seed(i)
for(i in seq(10000)){
 
  if(sum(Craps(10,i)) == 10){ ##Alter seed value of 10 game sessions.
    print(i)
    break
  } else {i <- i+1} ## Add to seed value and keep trying.
}
Craps(10,seed= i)
```

Yeah! With "1" denoting "Win", we just won 10 games!

**3)**
```{r eval=FALSE}
setwd("/Users/michaeltackerman/Documents/Bios6301/football-values-master/2015")

ffvalues <- function(path, file='outfile.csv', nTeams=12, cap=200,
                     posReq=c(qb=1, rb=2, wr=3, te=1, k=1),
                     points=c(fg=4, xpt=1, pass_yds=1/25, pass_tds=4, 
                     pass_ints=-2, rush_yds=1/10, rush_tds=6, fumbles=-2, 
                     rec_yds=1/20, rec_tds=6)) {
  
qb <- read.csv("proj_qb15.csv")  ## read in CSV files
rb <- read.csv("proj_rb15.csv")
te <- read.csv("proj_te15.csv")
wr <- read.csv("proj_wr15.csv")
k <- read.csv("proj_k15.csv")  

k[,'pos']  <- 'k'
qb[,'pos'] <- 'qb'
rb[,'pos'] <- 'rb'  
te[,'pos'] <- 'te'
wr[,'pos'] <- 'wr'

wr$points <- with(wr, rush_yds/10 + rush_tds*6 + rec_yds/20 + rec_tds*6 - 
                    fumbles*2)
te$points <- with(te, rec_yds/20 + rec_tds*6 - fumbles*2)
qb$points <- with(qb, 4*rush_tds+pass_yds/25+pass_tds*4+pass_ints*(-2)
                    +rush_yds/10+rush_tds*6+fumbles*(-2))
k$points <- with(k, 4*fg+xpt)
rb$points <- with(rb, rush_yds/10 + rush_tds*6 + rec_yds/20 + rec_tds*6 - 
                    fumbles*2)
k  <- k[,c('PlayerName','pos','points')]
qb <- qb[,c('PlayerName','pos','points')]
rb <- rb[,c('PlayerName','pos','points')] 
te <- te[,c('PlayerName','pos','points')]
wr <- wr[,c('PlayerName','pos','points')]

x <- rbind(k,qb,rb,te,wr)

x2 <- x[order(x[,'points'], decreasing=TRUE),]

k.ix <- which(x2[,'pos']=='k') 
qb.ix <- which(x2[,'pos']=='qb') 
rb.ix <- which(x2[,'pos']=='rb') 
te.ix <- which(x2[,'pos']=='te') 
wr.ix <- which(x2[,'pos']=='wr')

x2[k.ix, 'marg'] <- x2[k.ix,'points'] - x2[k.ix[12],'points']
x2[qb.ix, 'marg'] <- x2[qb.ix,'points'] - x2[qb.ix[12],'points']
x2[rb.ix, 'marg'] <- x2[rb.ix,'points'] - x2[rb.ix[24],'points']
x2[te.ix, 'marg'] <- x2[te.ix,'points'] - x2[te.ix[12],'points']
x2[wr.ix, 'marg'] <- x2[wr.ix,'points'] - x2[wr.ix[36],'points']

x3 <- x2[x2[,'marg'] >= 0,]

# re-order by marginal points
x3 <- x3[order(x3[,'marg'], decreasing=TRUE),]
rownames(x3) <- NULL

x3[,'value'] <- x3[,'marg']*(12*200-nrow(x3))/sum(x3[,'marg']) + 1

 write.csv(x3, file)

  return(x3)
}
###########################################################


# How many players are worth more than $20? (1 point)

x1 <- ffvalues(.)
sum(x1[,'value']>20)

## Who is 15th most valuable running back (rb)? (1 point)

x1[x1[,'pos']=='rb',][15,]

#  How many players are worth more than $20? (1 point)

x2 <- ffvalues(getwd(), '16team.csv', nTeams=16, cap=150)
sum(x2[,'value']>20)

#  1.  How many wide receivers (wr) are in the top 40? (1 point)

sum(x2[x2['pos']=='wr',]$value >40)
##############################
    x3 <- ffvalues('.', 'qbheavy.csv', posReq=c(qb=2, rb=2, wr=3, te=1, k=0),
            points=c(fg=0, xpt=0, pass_yds=1/25, pass_tds=6, pass_ints=-2,
                    rush_yds=1/10, rush_tds=6, fumbles=-2, rec_yds=1/20, rec_tds=6))


    #  How many players are worth more than $20? (1 point)
    
sum(x2[,'value']>20)

   # How many quarterbacks (qb) are in the top 30? (1 point)

sum(x2[1:30,'pos']=='qb')
```

**4)**
```{r}
objs <- mget(ls("package:base"), inherits = TRUE)
funs <- Filter(is.function, objs)
## 1)
argLengths = lapply(funs, function(f) length(formals(f)))

funs[[which.max(argLengths)]]

## 2)
sum(argLengths == 0)
```

