---
title: "Stat Computing HW1"
author: "Michael Ackerman"
date: "September 8th"
output: pdf_document
---

```{r}
gender <- c('M','M','F','M','F','F','M','F','M')
age <- c(34, 64, 38, 63, 40, 73, 27, 51, 47)
smoker <- c('no','yes','no','no','yes','no','no','no','yes')
exercise <- factor(c('moderate','frequent','some','some','moderate','none','none','moderate','moderate'),
                    levels=c('none','some','moderate','frequent'), ordered=TRUE
)
los <- c(4,8,1,10,6,3,9,4,8)
x <- data.frame(gender, age, smoker, exercise, los)
x
```

Model:

```{r}
model <- lm(los ~ gender + age + smoker + exercise, dat=x)
```

\textbf{1)} It appears that has being a male has the greatest effect on your length of stay with a value of 4.509.

\textbf{2)}
```{r}
mod <- lm(los~gender)
summary(mod)
```

\textbf{3)}
```{r}
mod$coef
```
\textbf{4)}
```{r}
sqrt(diag(vcov(mod)))
```
\textbf{5)}
```{r}
mod <- lm(los ~ gender, dat=x)
mod.c <- coef(summary(mod))
teststat <- mod.c[,1]/mod.c[,2]
2*pt(teststat['genderM'], 7, lower.tail = FALSE)
```
\textbf{6)}
```{r}
fitted(mod)-predict(mod)
```
It looks like the functions work a tiny bit different, yet create essentially the same vector. 
\textbf{7)}
```{r}
newdata <- data.frame(gender=c('F','M','F'))
predict(mod,newdata)
```
\textbf{8)}
```{r}
residuals <- fitted(mod)- x$los
residuals
```
\textbf{9)}
```{r}
residuals(mod)
```
Same result in absolute value. 

\textbf{10)}
```{r}
sum(residuals^2) - deviance(mod)
```
These functions do the same thing.

\textbf{11)}
```{r}
sqrt(deviance(mod)/df.residual(mod))
```
This is the same as the Residual Standard Error shown by the original model.

\textbf{12)}
```{r}
xmen <- x[gender== "M",]
women <- x[gender=="F",]
```
X-Men is the proper labeling. 

\textbf{13)}
```{r}
var(xmen$los)
var(women$los)
```
\textbf{14)}
```{r}
t.test(women$los, xmen$los)
t.test(women$los, xmen$los, var.equal = T)

```

The equal variance assumption causes that p-value in the t-test comparison to be the same as the  full model with the genderM variable.

```{r}
##this is nice and consice. 
t.test(los ~ gender, dat=x, var.equal=TRUE)
```