---
title: "Stat Comp Final"
author: "Michael Ackerman"
date: "November 23, 2015"
output: pdf_document
---

1. Read and combine the projection data (five files) into one data set, adding a position column.

1. The NFL season is 17 weeks long, and 10 weeks have been completed.  Each team plays 16 games and has one week off, called the bye week.  Four teams have yet to have their bye week: CLE, NO, NYG, PIT.  These four teams have played ten games, and every other team has played nine games.  Multiply the numeric columns in the projection data by the percentage of games played (for example, 10/16 if team is PIT).

1. Sort and order the data by the `fpts` column descendingly.  Subset the data by keeping the top 20 kickers, top 20 quarterbacks, top 40 running backs, top 60 wide recievers, and top 20 tight ends.  Thus the projection data should only have 160 rows.

1. Read in the observed data (`nfl_current15.csv`)

1. Merge the projected data with the observed data by the player's name. Keep all 160 rows from the projection data. If observed data is missing, set it to zero.

You can directly compare the projected and observed data for each player.  There are fifteen columns of interest.

1. Take the difference between the observed data and the projected data for each category. Split the data by position, and keep the columns of interest.

You will now have a list with five elements.  Each element will be a matrix or data.frame with 15 columns.

```{r question3}
path = "~/Documents/football-values-master-2/2015/"
file='outfile.csv'
  k <- read.csv(file.path(path, 'proj_k15.csv'), header=TRUE, stringsAsFactors=FALSE)
  qb <- read.csv(file.path(path, 'proj_qb15.csv'), header=TRUE, stringsAsFactors=FALSE)
  rb <- read.csv(file.path(path, 'proj_rb15.csv'), header=TRUE, stringsAsFactors=FALSE)
  te <- read.csv(file.path(path, 'proj_te15.csv'), header=TRUE, stringsAsFactors=FALSE)
  wr <- read.csv(file.path(path, 'proj_wr15.csv'), header=TRUE, stringsAsFactors=FALSE)
observed <- read.csv(file.path(path, 'nfl_current15.csv'), header=TRUE, stringsAsFactors=FALSE)
  # generate unique list of column names
  cols <- unique(c(names(k), names(qb), names(rb), names(te), names(wr)))
  k[,'pos'] <- 'k'
  qb[,'pos'] <- 'qb'
  rb[,'pos'] <- 'rb'
  te[,'pos'] <- 'te'
  wr[,'pos'] <- 'wr'
  cols <- c(cols, 'pos')
  # create common columns in each data.frame
  # initialize values to zero
  k[,setdiff(cols, names(k))] <- 0
  qb[,setdiff(cols, names(qb))] <- 0
  rb[,setdiff(cols, names(rb))] <- 0
  te[,setdiff(cols, names(te))] <- 0
  wr[,setdiff(cols, names(wr))] <- 0
  # combine data.frames by row, using consistent column order
  x <- rbind(k[,cols], qb[,cols], rb[,cols], te[,cols], wr[,cols])
  # furthermore, gsub does global replacement, not single replacement
  names(x) <- gsub('[.]', '', names(x))
  
  # bye-ajusted values
  x[,c(-1:-2,-19)] <- (9/16)*x[,c(-1:-2,-19)]
  nobye <- which(x[,'Team'] == c('CLE', 'NO', 'NYG', 'PIT'))
  x[nobye, c(-1:-2,-19)]  <- (10/9)*x[nobye,c(-1:-2,-19)]

x <- x[order(x$fpts, decreasing=T),]
posis <- unique(x$pos)
numb <- c(20,40,60,20,20) #amount of each position we want

y <- NULL # Remove the extra players we dont care to look at
for(i in seq_along(posis)){
y <- rbind(y, x[which(x$pos==posis[i])[1:numb[i]],])
}

# Merge Observed and predicted data
x <- y[order(y$fpts, decreasing=T),]
pro <- x
df <- merge(x, observed, by.x = "PlayerName", by.y = "Name", all.x = T, all.y =F)
df <- df[order(df$fpts, decreasing=T),]
rownames(df) <- NULL
df[is.na(df)] <- 0

name <- c('field goals','field goals attempted','extra points','passing attempts','passing completions', 'passing yards','passing touchdowns','passing interceptions','rushing attempts','rushing yards', 'rushing touchdowns','receiving attempts','receiving yards','receiving touchdowns','fumbles')

projected_col=c('fg','fga','xpt','pass_att','pass_cmp','pass_yds','pass_tds','pass_ints',
                           'rush_att','rush_yds','rush_tds','rec_att','rec_yds','rec_tds','fumbles')
observed_col=c("FGM","FGA","XPM","Att.pass","Cmp.pass","Yds.pass","TD.pass","Int.pass",
                           "Att.rush","Yds.rush","TD.rush","Rec.catch","Yds.catch","TD.catch","Fmb")

diff <- cbind( df[,'pos'], df[,observed_col]-df[,projected_col] )
names(diff) <- c('pos', name) 

ldiff <- split(diff[,2:16], diff[,'pos']) #list of matrices by 'pos'

```

Task 2: Creating League S3 Class (80 points)
------------

Create an S3 class called `league`.  Place all code at the end of the instructions.

1. Create a function `league` that takes 5 arguments (`stats`, `nTeams`, `cap`, `posReq`, `points`).  It should return an object of type `league`.  Note that all arguments should remain attributes of the object.  They define the league setup and will be needed to calculate points and dollar values.

1. Create a function `calcPoints` that takes 1 argument, a league object.  It will modify the league object by calculating the number of points each player earns, based on the league setup.

1. Create a function `buildValues` that takes 1 argument, a league object.  It will modify the league object by calculating the dollar value of each player.

    As an example if a league has ten teams and requires one kicker, the tenth best kicker should be worth $1.  All kickers with points less than the 10th kicker should have dollar values of $0.

1. Create a `print` method for the league class. It should print the players and dollar values (you may choose to only include players with values greater than $0).
1. Create a `plot` method for the league class. Add minimal plotting decorations (such as axis labels).


```{r}
#league function will create a list where the first element are the stats and proceeding elements are the characteristcs. it will define this list as the class(list) <- 'league'

#the following functions will pull the elements of that list and computed the desired values, adding them to the first data frame and returning another list in the league class

league <- function(stats, nTeams=12, cap=200,posReq=c(qb=1, rb=2, wr=3, te=1, k=1),
                     points=c(fg=4, xpt=1, pass_yds=1/25, pass_tds=4, pass_ints=-2,
                     rush_yds=1/10, rush_tds=6, fumbles=-2, rec_yds=1/20, rec_tds=6)){
 if(is.list(posReq)) posReq <- unlist(posReq)
 if(is.list(points)) points <- unlist(points)
  info <- list(stats=stats, nTeams=nTeams, cap=cap, posReq=posReq, points=points)
  class(info) <- 'league'
  return(info)
}
##############################################################

calcPoints<- function(p){UseMethod("calcPoints")}
buildValues <- function(p){UseMethod("buildValues")}

calcPoints.league <- function(p){ #calculate point
  x <- y <- p[[1]]

  for(i in names(p$points)) {
    x[,sprintf("p_%s", i)] <- x[,i]*p$points[[i]]    }
  
  y[,'points'] <- rowSums(x[,grep("^p_", names(x))])
  p[[1]] <- y
  return(p)
}
################################################################

buildValues.league <- function(stats){ #calculate dollar values
  p<- calcPoints(stats)
  x<- p[[1]]
  x2 <- x[order(x[,'points'], decreasing=TRUE),]
  # calculate marginal points by subtracting "baseline" player's points
  for(i in names(p$posReq)) {
    ix <- which(x2[,'pos'] == i)
    baseline <- p$posReq[[i]]*p$nTeams
    if(baseline == 0) {
      x2[ix, 'marg'] <- -1
    } else {
      x2[ix, 'marg'] <- x2[ix,'points'] - x2[ix[baseline],'points']
    }
  }
  x3 <- x2
  x3[,'value'] <-  ifelse( x2[,'marg'] >= 0,
   x3[,'marg']*(p$nTeams*p$cap-nrow(x3))/sum(x3[which(x3[,'marg']>0),'marg']) +1, 0)
  x3[is.na(x3[,'value']),] <- 0
  
  x3[,'marg'] <- NULL
  x4 <- x3[rownames(x), ]
  p[[1]]<-x4
  return(p)
}
#p <-buildValues(l)
###############################################
print.league <- function(stats){
  x <-buildValues(stats)$stats
  x <- x[x[,'value']>0,]
  cat(paste("The",x$pos,  x$PlayerName, "is worth: $", round(x$value,2), "\n"))
}
###############################################
plot.league <- function(p){
  x <- buildValues(p)$stats
    x <- x[x[,'value']>0,]
  plot(x$value, ylab="Dollar Value", xlab="Ranking", main="Ranking to Player Value")
}
###############################################
hist.league <- function(p){
  x <- buildValues(p)$stats
  x <- x[x[,'value']>0,]
  hist(x$value, xlab= "Dollar Value", main= "Distribution of Player Value") 
}
###############################################
boxplot.league <- function(p){
  x <- buildValues(p)$stats
  x <- x[x[,'value']>0,c('pos','value')]
  boxplot(value~pos, data= x) 
}
#rm(print.league)
``` 

```{r}
pos <- list(qb=1, rb=2, wr=3, te=1, k=1)
pnts <- list(fg=4, xpt=1, pass_yds=1/25, pass_tds=4, pass_ints=-2,
             rush_yds=1/10, rush_tds=6, fumbles=-2, rec_yds=1/20, rec_tds=6)
l <- league(stats=df, nTeams=10, cap=200, posReq=pos, points=pnts)

plot(l)
hist(l)
boxplot(l)
```

Task 3: Simulations with Residuals (40 points)
------------

Using residuals from task 1, create a list of league simulations.  The simulations will be used to generate confidence intervals for player values.  Place all code at the end of the instructions.

1. Create a function `addNoise` that takes 4 arguments: a league object, a list of residuals, number of simulations to generate, and a RNG seed. It will modify the league object by adding a new element `sims`, a matrix of simulated dollar values.

    The original league object contains a `stats` attribute. Each simulation will modify this by adding residual values. This modified `stats` data.frame will then be used to create a new league object (one for each simulation).  Calculate dollar values for each simulation.  Thus if 1000 simulations are requested, each player will have 1000 dollar values.  Create a matrix of these simulated dollar values and attach it to the original league object.

    As an example assume you want to simulate new projections for quarterbacks. The residuals for quarterbacks is a 20x15 matrix. Each row from this matrix is no longer identified with a particular player, but rather it's potential error. Given the original projection for the first quarterback, sample one value between 1 and 20. Add the 15 columns from the sampled row to the 15 columns for the first quarterback. Repeat the process for every quarterback. Note that stats can't be negative so replace any negative values with 0.

1. Create a `quantile` method for the league class; it takes at least two arguments, a league object and a probs vector. This method requires the `sims` element; it should fail if `sims` is not found. The `probs` vector should default to `c(0.25, 0.5, 0.75)`.  It should run `quantile` on the dollar values for each player.

1. Create a function `conf.interval`; it takes at least two arguments, a league object and a probs vector. This method requires the `sims` element; it should fail if `sims` is not found. It should return a new object of type `league.conf.interval`.

    The new object will contain the output of `quantile`. However, results should be split by position and ordered by the last column (which should be the highest probability) descendingly.  Restrict the number of rows to the number of required players at each position.

1. Create a `plot` method for the league.conf.interval class; it takes at least two arguments, a league.conf.interval object and a position. Plot lines for each probability; using the defaults, you would have three lines (0.25, 0.5, 0.75). Add minimal plotting decorations and a legend to distinguish each line.
```{r}
simValue1 <- function(...){UseMethod("simValue1")}

simValue1.league <- function(stats,resid){
  x<- stats$stats
  y <- x[,c('PlayerName', 'fpts')] #use later to reorder
  lx <- split(x, x[,'pos']) # yes thats a "L"
    
  lnew <- list()
  new1 <- NULL
  nums <- sapply(x, is.numeric)
  
  for(i in 1:5){
  posis <- names(lx[i])

 rowi <- floor(runif(dim(lx[[i]])[1], min = 1, max= dim(resid[[posis]])[1] ))

 x2 <- lx[[i]][,c("fg","fga","xpt","pass_att","pass_cmp","pass_yds","pass_tds","pass_ints","rush_att","rush_yds", "rush_tds", "rec_att", "rec_yds", "rec_tds", "fumbles")]
  
  lnew[[i]] <- cbind(lx[[i]][,c('PlayerName','pos')], x2+resid[[posis]][rowi,])
  colnames(lnew[[i]]) <- c('PlayerName','pos', colnames(x2))
  
  new1 <- rbind(new1 ,lnew[[i]])
  }

 new1[new1<0] <- 0
new1 <-  merge(y,new1, by = "PlayerName", all.x= FALSE,all.y=T, sort=FALSE)#restore original order
stats$stats <- new1

 val <-buildValues(stats)[[1]]$value
return(val)
  }
#####################################################
#simValue1(l, ldiff)

addNoise <- function(p,...){UseMethod("addNoise")}
addNoise.league <- function(stats, resid= ldiff, nsims=100, seed=120, warned = FALSE){
  if(nsims>4000 & warned == 0){warning("This is going to take a little while. Are you sure you want to do this? Is it really necessary? Alright, change the fifth argument 'warned?' to TRUE and go get some coffee. 10,000 simulations will take just under 5 minutes"); break}
  set.seed(seed)
  simulations <- matrix(NA, nrow=length(simValue1(stats, resid)) , ncol= nsims) 
  for(i in 1:nsims){
  simulations[,i] <- simValue1(stats, resid)
  }
stats$simulations <- simulations
return(stats)
}
######################################################
#j <- addNoise(stats, ldiff, 100,,T)

quantile.league <- function(stats, probs= c(0.25, 0.5, 0.75),resid= ldiff, nsims=100){
  if(is.null(stats$simulations)){
    stats = addNoise(stats,resid, nsims)
    message(paste("This is a simulation calculated by this function becuase you didn't provide one yourself. Your account has been automatically charged $6 for this extra work. Next time you should atleast declare an amount of simulations as the fourth argument ...but you will be charged again."))
  }
  quant1 <- apply(stats$simulations, 1, quantile, probs)
  quant <- quant1[,which(quant1[2,]>0)] #Only the good players
 return(quant1)
}
######################################################
#quantile(l,probs= c(0.25, 0.5, 0.75))

conf.interval <- function(...){UseMethod("conf.interval")}

conf.interval.league <- function(stats, probs= c(0.25, 0.5, 0.75),resid= ldiff, nsims=100){
  #quantile(stats,probs= c(0.25, 0.5, 0.75),resid= ldiff, nsims=100) 
  x <- stats$stats
  q <- quantile(stats)
  q <- t(q)
  
  H<- data.frame(PlayerName=x$PlayerName, pos=x$pos, lower=q[,1], mid=q[,2], upper=q[,3])
  H <- H[order(H[,'upper'], decreasing=TRUE),]
H <- split(H, H$pos)
n <- stats$nTeams*stats$posReq
for(i in names(n)){
  H[[i]] <- H[[i]][1:n[[i]],]
  rownames(H[[i]])<- NULL
}
Conf.int <- H
  class(Conf.int) <- 'league.conf.interval'
return(Conf.int)
}
######################################################
#conf.interval(l)

plot.league.conf.interval <- function(intervals, pos="qb"){
  d <- intervals[[pos]]
  plot(d$m, ylab= "Value", xlab= "Per Player")
  lines(d$l)
  lines(d$u)
}

```

```{r}
l1 <- addNoise(l, ldiff, 1000)
#quantile(l1)
(ci <- conf.interval(l1) )
plot(ci, 'qb')
plot(ci, 'rb')
plot(ci, 'wr')
plot(ci, 'te')
plot(ci, 'k')
```

